---
title: "DS606-Lab2"
author: "George Cruz"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

```{r}
ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram()

ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 15)

ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 150)
```

### Exercise 1

It appears that some flights are leaving before their departure time and this is not apparent in the first histogram with the lower binwidth. 

```{r code-chunk-label}
lax_flights <- nycflights %>%
  filter(dest == "LAX")
ggplot(data = lax_flights, aes(x = dep_delay)) +
  geom_histogram()

lax_flights %>%
  summarise(mean_dd   = mean(dep_delay), 
            median_dd = median(dep_delay), 
            n         = n())
```

### Exercise 2

```{r}
sfo_feb_flights <- nycflights %>%
  filter(dest == "SFO", month == 2)
sfo_feb_flights %>% summarise(count = n())
```
68 flights departed to San Francisco in February.

### Exercise 3
```{r}
ggplot(data = sfo_feb_flights, aes(x = arr_delay)) +
  geom_histogram()

sfo_feb_flights %>%
  summarise(median_ad = median(arr_delay), iqr_ad = IQR(arr_delay), n_flights = n())
```
The distribution is right-skewed and unimodal.  The median and IQR would be the best summary statistics to avoid variability due to the few outliers. The Median arrival delay is -11 and the IQR is 23.25

### Exercise 4
```{r}
sfo_feb_flights %>%
  group_by(carrier) %>%
  summarise(median_ad = median(arr_delay), iqr_ad = IQR(arr_delay), n_flights = n())
```
Delta Airlines (DL) and United Airlines (UA) have the most variability in arrival delays. 

### Exercise 5
```{r}
nycflights %>%
  group_by(month) %>%
  summarise(mean_dd = mean(dep_delay)) %>%
  arrange(desc(mean_dd))
```
```{r}
nycflights %>%
  group_by(month) %>%
  summarise(median_dd = median(dep_delay)) %>%
  arrange(desc(median_dd))
```
The Median is a more robust value as it is less susceptible to outliers.  In this case, I think measuring the outliers and thus, using the mean, would give us a better chance to avoid most delays. 

### Exercise 6
```{r}
nycflights <- nycflights %>%
  mutate(dep_type = ifelse(dep_delay < 5, "on time", "delayed"))

nycflights %>%
  group_by(origin) %>%
  summarise(ot_dep_rate = sum(dep_type == "on time") / n()) %>%
  arrange(desc(ot_dep_rate))
```

Based only On Time Departure Percentage, La Guardia would be the logical choice.

```{r}
ggplot(data = nycflights, aes(x = origin, fill = dep_type)) +
  geom_bar()
```

### Exercise 7
```{r}
nycflights <- nycflights %>%
  mutate(avg_speed = distance/(air_time/60))
head(nycflights)
```


### Exercise 8
```{r}
ggplot(nycflights, aes(distance, avg_speed )) +
        geom_point(size=1, shape=15)
```

There appears to be a non-linear relationship between avg_speed and distance.  There is an increase of avg speed when distance increases.  


### Exercise 9
```{r}
airlines <- c("AA", "DL", "UA")
selectedFlights <- nycflights %>%
  filter(carrier %in% airlines)

qplot(x = dep_delay, y = arr_delay, data = selectedFlights, color = carrier)
```

George Cruz 