---
title: "DS606-Lab7"
author: "George Cruz"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

#Inference for Numerical Data
```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(infer)
data(yrbss)

```


### Exercise 1
**What are the cases in this data set? How many cases are there in our sample?**
Individual Youth behavior.  There are 13,583 cases in our sample. Each case represents an individual student's behavior and demographic info.

```{r code-chunk-label}
glimpse(yrbss)
```

```{r}
summary(yrbss$weight)
```

### Exercise 2
**How many observations are we missing weights from?**
We're missing 1,004.


```{r}
yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(yrbss$physically_active_7d > 2, "yes", "no"))
```

### Excercise 3
**Make a side-by-side boxplot of physical_3plus and weight. Is there a relationship between these two variables? What did you expect and why?**

```{r}
boxplot(yrbss$weight ~ yrbss$physical_3plus, col="orange", main="Distribution of Physical Activity and Weight", ylab="Weight", xlab="Physical Activity")
```

There appears to be minimal change based on the physical activity.  The mean weight of the "yes" is slightly higher than the "no" and it appears to be a more normalized distribution.  I expected the mean weight of the "no" to be higher based on the physical activity. 

```{r}
yrbss %>%
  group_by(physical_3plus) %>%
  summarise(mean_weight = mean(weight, na.rm = TRUE))
```

###Exercise 4
The conditions we need for inference on one proportion are:
Random: The data comes from a random or randomized experiment.
Normal: The sampling distribution of p^ needs to be approximately normal — needs at least 10 expected successes and 10 expected failures.
Independent: Individual observations need to be independent. 

###Exercise 5
H0: There's no statistical difference in weight due to exercise. 
H1: The average weights are different between those who exercise at least 3 times a week. 


```{r}
obs_diff <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

null_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()
```

### Exercise 6

```{r}
null_dist %>%
  filter(stat >= obs_diff)
```
We found no permutation with a difference with at least obs_stat.

```{r}
null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")
```

What this warning means is that, although the result in calculations is 0, p-value is an *approximation* and not the real zero.  As the number of repetitions grow, the p-value becomes increasingly small or approaches 0.


### Exercise 7

```{r}

p1 <- yrbss %>%
    filter(physical_3plus == "yes") %>%
    summarise(mean_weight = mean(weight, na.rm = TRUE), 
              total = n())
p2 <- yrbss %>%
  filter(physical_3plus == "no") %>%
    summarise(mean_weight = mean(weight, na.rm = TRUE),
              total = n())

n <- 13583 - 1004

prop1 <- p1$total/n
prop2 <- p2$total/n

mean_diff <- p1$mean_weight - p2$mean_weight

se <- ( ((prop1 * (1 - prop1)) / n) +  ((prop2 * (1 - prop2)) / n)) ** 0.5
z <- 1.96

me <-z * se

ci_95 <- c(mean_diff - me, mean_diff + me)

ci_95

```
The 95% confidence interval of the difference between those who exercise regularly and those who do not is between 1.76 and 1.79kg.  This means that there is a difference between those who exercise and those who do not and we should reject the null hypothesis.

### Exercise 8

```{r}

z = 1.96
mean_height <- mean(yrbss$height, na.rm = TRUE)
sd_height <- sd(yrbss$height, na.rm = TRUE)
total_height <- yrbss %>% 
  filter(!is.na(height)) %>%
  summarise(n = n())

total_height <- total_height$n

ci_low <- mean_height - (z * (sd_height/sqrt(total_height)))
ci_high <- mean_height + (z * (sd_height/sqrt(total_height)))

ci_95 <- c(ci_low, ci_high)
ci_95
```

The 95% confidence interval is between 1.6894 and 1.6930

### Exercise 9

```{r}
z = 1.64
#mean_height <- mean(yrbss$height, na.rm = TRUE)
#sd_height <- sd(yrbss$height, na.rm = TRUE)
#total_height <- yrbss %>% 
#  filter(!is.na(height)) %>%
#  summarise(n = n())

#total_height <- total_height$n

ci_low <- mean_height - (z * (sd_height/sqrt(total_height)))
ci_high <- mean_height + (z * (sd_height/sqrt(total_height)))

ci_90 <- c(ci_low, ci_high)
ci_90
```

The 90% confidence interval is more narrow: 1.6897 and 1.6927

### Exercise 10

H0 = There's no statistical difference in height between those who exercise regularly and those who don't.

H1 = There is a difference in height between the two groups. 


```{r}

boxplot(yrbss$height ~ yrbss$physical_3plus, col="orange", main="Distribution of Physical Activity and Height", ylab="Height", xlab="Physical Activity")
```

There appears to be a difference in height between the two groups. 
```{r}
yrbss %>%
  group_by(physical_3plus) %>%
  summarise(mean_height = mean(height, na.rm = TRUE))
```

```{r}
obs_diff_ht <- yrbss %>%
  specify(height ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

null_dist_ht <- yrbss %>%
  specify(height ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

null_dist_ht %>%
  get_p_value(obs_stat = obs_diff_ht, direction = "two_sided")

```


What this warning means is that, although the result in calculations is 0, p-value is an *approximation* and not the real zero.  As the number of repetitions grow, the p-value becomes increasingly small or approaches 0.

```{r}

p1 <- yrbss %>%
    filter(physical_3plus == "yes") %>%
    summarise(mean_height = mean(height, na.rm = TRUE), 
              total = n())
p2 <- yrbss %>%
  filter(physical_3plus == "no") %>%
    summarise(mean_height = mean(height, na.rm = TRUE),
              total = n())

n <- 12364

prop1 <- p1$total/n
prop2 <- p2$total/n

mean_diff <- p1$mean_height - p2$mean_height

se <- ( ((prop1 * (1 - prop1)) / n) +  ((prop2 * (1 - prop2)) / n)) ** 0.5
z <- 1.96

me <-z * se

ci_95 <- c(mean_diff - me, mean_diff + me)

ci_95
```

The 95% confidence interval of the difference between those who exercise regularly and those who do not is between 0.02 and 0.05.  This means that there is a difference between those who exercise and those who do not and we should reject the null hypothesis.


### Exercise 11

```{r}

yrbss %>% 
  group_by(hours_tv_per_school_day) %>% 
  summarise(n = n(), .groups="drop_last")

```

7 options, unless we count NA, then 8.


### Exercise 12

Research question: Is there evidence that students weight is affected by the sleep time?

H0: There is no relationship between weight and sleep for students.
H1: There is a relationship between weight and sleep. 

α Level: .05 (95% confident)


```{r}

yrbss <- yrbss %>% 
  mutate(sleep_7plus = ifelse(yrbss$school_night_hours_sleep > 6, "yes", "no"))

boxplot(yrbss$weight ~ yrbss$sleep_7plus, col="orange", main="Distribution of Good night sleep and Weight", ylab="Weight", xlab="Good night sleep")
```

The boxplot points to a slight difference in weight.

```{r}
weight_no <- yrbss %>% 
    filter(sleep_7plus == "no") %>%
    summarise(w_avg = mean(weight, na.rm = TRUE), 
              w_sd = sd(weight, na.rm=TRUE),
              total = n())

weight_yes <- yrbss %>% 
    filter(sleep_7plus == "yes") %>%
    summarise(w_avg = mean(weight, na.rm = TRUE), 
              w_sd = sd(weight, na.rm=TRUE),
              total = n())

mean_diff <- weight_no$w_avg - weight_yes$w_avg

se <- 
  sqrt(
  ((weight_no$w_avg**2) / weight_no$total) +
  ((weight_yes$w_avg**2) / weight_yes$total))

ci_low <- mean_diff - 1.96 * se
ci_high <- mean_diff + 1.96 * se

ci_95 <- c(ci_low, ci_high)

ci_95
```

Because the CI is between -0.92 and 3.92 we will accept the null hypothesis that there is no relationship between weight and sleep.  These results are somewhat surprising. 
...


