---
title: "Lab 3"
author: "George Cruz"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1

A streak of 1 means 1 hit.  A streak of zero means a miss.


### Exercise 2
```{r}
kobe_streak <- calc_streak(kobe_basket$shot)
ggplot(data = kobe_streak, aes(x = length)) +
  geom_bar()
```
The distribution is right-skewed.  The typical streak length was 0 and the longest streak is 4.

### Exercise 3
```{r}
coin_outcomes <- c("heads", "tails")
sample(coin_outcomes, size = 1, replace = TRUE)

sim_fair_coin <- sample(coin_outcomes, size = 100, replace = TRUE)

sim_fair_coin
table(sim_fair_coin)


```
```{r}
set.seed(989) 
sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE, 
                          prob = c(0.2, 0.8))
table(sim_unfair_coin)
```
Only 11 times did the unfair coin come up heads. 

### Exercise 4
What change needs to be made to the sample function so that it reflects a shooting percentage of 45%? Make this adjustment, then run a simulation to sample 133 shots. Assign the output of this simulation to a new object called sim_basket.

```{r}
set.seed(100)
shot_outcomes <- c("H", "M")
sim_basket <- sample(shot_outcomes, size = 133, replace = TRUE, prob=c(0.45, 0.55))
table(sim_basket)
```
I added the seed so the values did not change.

### Exercise 5
```{r}
sim_streak <- calc_streak(sim_basket)

```

### Exercise 6
```{r}
ggplot(data = sim_streak, aes(x = length)) +
  geom_bar()
```

Distribution is right-skewed, longest streak was 6 and typical streak was 0. 

### Exercise 7
I would expect a similar distribution. 

### Exercise 8 
The distributions are somewhat similar between Kobe and the simulation.  The longest streak is somewhat different which lets me to believe Kobe doesn't fit the hot hand model. 




