---
title: "Lab 9 - Multiple linear regression"
author: "George Cruz"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

# Multiple linear regression

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(GGally)
```

### Exercise 1

**Is this an observational study or an experiment? The original research question posed in the paper is whether beauty leads directly to the differences in course evaluations. Given the study design, is it possible to answer this question as it is phrased? If not, rephrase the question.**

This is an observational study.  Becauseof this, we cannot prove causation.  Therefore, a better question to ask would be: *Is there a correlation between beauty and course evaluations?*

### Exercise 2
**Describe the distribution of score. Is the distribution skewed? What does that tell you about how students rate courses? Is this what you expected to see? Why, or why not?**

```{r, warning=FALSE, message=FALSE}
ggplot(evals, aes(x=score)) + 
  geom_histogram()
```

The distribution is left-skewed.  It seems to indicate that students tend to rate courses high.  This is what I expected to see. 

### Exercise 3

**Excluding score, select two other variables and describe their relationship with each other using an appropriate visualization.**

```{r, error=FALSE, warning=FALSE, message=FALSE}
ggplot(data = evals, aes(x = cls_perc_eval, y = bty_avg)) +
  geom_point()+
  geom_smooth(method="lm", se=TRUE)
```
## Simple linear regression

The fundamental phenomenon suggested by the study is that better looking teachers are evaluated more favorably. Let’s create a scatterplot to see if this appears to be the case:

```{r}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_point()
```

### Exercise 4
**Replot the scatterplot, but this time use geom_jitter as your layer. What was misleading about the initial scatterplot?**

```{r}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter()
```

There appears to be more points than in the first scatterplot and, we can also see some correlation or trend where before we could not. 

### Exercise 5 

**Let’s see if the apparent trend in the plot is something more than natural variation. Fit a linear model called m_bty to predict average professor score by average beauty rating. Write out the equation for the linear model and interpret the slope. Is average beauty score a statistically significant predictor? Does it appear to be a practically significant predictor?**

```{r}
m_bty <- lm( score ~ bty_avg, evals)
summary(m_bty)
```

$y = b_0 + b_1 * bty\_avg$

$y = 3.88034 + 0.06664 * bty\_avg$

The slope means that for every point increase in bty_avg, we will see a 0.067 increase in score.  Based on this slope and the adjusted $R^2$, Although it is a **statistically significant predictor**, I would say this is not a practically significant
predictor.

```{r}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE)
```

### Exercise 6
**Use residual plots to evaluate whether the conditions of least squares regression are reasonable. Provide plots and comments for each one (see the Simple Regression Lab for a reminder of how to make these).**

*Nearly Normal Residuals*

```{r}
ggplot(data = m_bty, aes(x = .resid)) +
  geom_histogram(binwidth = 0.25) +
  xlab("Residuals")
```

We have somewhat nearly normal residuals. 

*Constant variability*

```{r}
ggplot(m_bty, aes(x = .fitted, y = .resid)) + 
  geom_point() +
  geom_hline(yintercept = 0, linetype=1, color="blue") 
```

The variability of the numbers does not appear constant. 

#Multiple linear regression
The data set contains several variables on the beauty score of the professor: individual ratings from each of the six students who were asked to score the physical appearance of the professors and the average of these six scores. Let’s take a look at the relationship between one of these scores and the average beauty score.

```{r}
ggplot(data = evals, aes(x = bty_f1lower, y = bty_avg)) +
  geom_point()
evals %>% 
  summarise(cor(bty_avg, bty_f1lower))
```

As expected, the relationship is quite strong—after all, the average score is calculated using the individual scores. You can actually look at the relationships between all beauty variables (columns 13 through 19) using the following command:
```{r}
evals %>%
  select(contains("bty")) %>%
  ggpairs()
```

These variables are collinear (correlated), and adding more than one of these variables to the model would not add much value to the model. In this application and with these highly-correlated predictors, it is reasonable to use the average beauty score as the single representative of these variables.

In order to see if beauty is still a significant predictor of professor score after you’ve accounted for the professor’s gender, you can add the gender term into the model.

```{r}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```

### Exercise 7
P-values and parameter estimates should only be trusted if the conditions for the regression are reasonable. Verify that the conditions for this model are reasonable using diagnostic plots.

```{r, warning=FALSE, message=FALSE}
library(cowplot)
p1 <- ggplot(m_bty_gen, aes(sample = .resid)) + 
  stat_qq() + 
  stat_qq_line()

p2 <- ggplot(data = m_bty_gen, aes(x = .resid)) +
  geom_histogram(binwidth = 0.25) +
  xlab("Residuals")

plot_grid(p1, p2, labels="AUTO")
```

Residuals appear nearly normal. 

```{r}
ggplot(m_bty_gen, aes(x = .fitted, y = .resid)) + 
  geom_point() +
  geom_hline(yintercept = 0, linetype=1, color="blue") 
```

I don't think I see a constant variability of residuals but they are spread without an apparent pattern. 

### Exercise 8

**Is bty_avg still a significant predictor of score? Has the addition of gender to the model changed the parameter estimate for bty_avg?**

Yes.  Adding gender has increased the parameter estimate for beaty.

### Exercise 9
**What is the equation of the line corresponding to those with color pictures? (Hint: For those with color pictures, the parameter estimate is multiplied by 1.) For two professors who received the same beauty rating, which color picture tends to have the higher course evaluation score?**

We don't have any linear regression using color pictures. If we performed a linear regression adding the pic_color it would be: 

```{r}
m_bty_pic <- lm(score ~ bty_avg + pic_color, data = evals)
summary(m_bty_pic)
```

$score= 4.06318 + 0.05548 * bty\_avg - 0.16059$

A black and White picture will give out a greater score than a color one. 

### Exercise 10
**Create a new model called m_bty_rank with gender removed and rank added in. How does R appear to handle categorical variables that have more than two levels? Note that the rank variable has three levels: teaching, tenure track, tenured.**

```{r}
m_bty_rank <- lm(score ~ bty_avg + rank, data = evals)
summary(m_bty_rank)
```

It appears that R sets one level as the 0 value and then calculates slopes for the other levels.

# The search for the best model 

We will start with a full model that predicts professor score based on rank, gender, ethnicity, language of the university where they got their degree, age, proportion of students that filled out evaluations, class size, course level, number of professors, number of credits, average beauty rating, outfit, and picture color.


### Exercise 11
**Which variable would you expect to have the highest p-value in this model? Why? Hint: Think about which variable would you expect to not have any association with the professor score.**

cls_credits seems to be the one. 


Let’s run the model…

```{r}
m_full <- lm(score ~ rank + gender + ethnicity + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

### Exercise 12
**Check your suspicions from the previous exercise. Include the model output in your response.**

I was incorrect.  The highest p-value is for cls_profs.

### Exercise 13
**Interpret the coefficient associated with the ethnicity variable.**

If all other factors are the same, the ethnicitynotminority will increase the score 0.1234929.

### Exercise 14
**Drop the variable with the highest p-value and re-fit the model. Did the coefficients and significance of the other explanatory variables change? (One of the things that makes multiple regression interesting is that coefficient estimates depend on the other variables that are included in the model.) If not, what does this say about whether or not the dropped variable was collinear with the other explanatory variables?**

```{r}
m_full2 <- lm(score ~ rank + gender + ethnicity + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full2)
```

By dropping this variable, the coefficients and the p-values changed slightly for the others. I would say the dropped variable was somewhat collinear with the others. 

### Exercise 15
**Using backward-selection and p-value as the selection criterion, determine the best model. You do not need to show all steps in your answer, just the output for the final model. Also, write out the linear model for predicting score based on the final model you settle on.**

```{r}
m_best <- lm(score ~ gender + ethnicity + language + age + cls_perc_eval 
             + cls_credits + bty_avg 
             + pic_color, data = evals)
summary(m_best)
```

$score = 3.771922 + 0.207112 * gendermale + 0.167872 * ethnicitynot minority - 0.206178 * languagenon-english - 0.006046 * age + 0.004656 * cls_perc_eval + 0.505306 * cls_creditsone credit + 0.051069 * bty_avg - 0.190579 * pic_colorcolor$

### Exercise 16
**Verify that the conditions for this model are reasonable using diagnostic plots.**

```{r, warning=FALSE, message=FALSE}
p1 <- ggplot(m_best, aes(sample = .resid)) + 
  stat_qq() + 
  stat_qq_line()

p2 <- ggplot(data = m_best, aes(x = .resid)) +
  geom_histogram(binwidth = 0.25) +
  xlab("Residuals")

plot_grid(p1, p2, labels="AUTO")
```

Residuals appear nearly normal. 

```{r}
ggplot(m_best, aes(x = .fitted, y = .resid)) + 
  geom_point() +
  geom_hline(yintercept = 0, linetype=1, color="blue") 
```



### Exercise 17
**The original paper describes how these data were gathered by taking a sample of professors from the University of Texas at Austin and including all courses that they have taught. Considering that each row represents a course, could this new information have an impact on any of the conditions of linear regression?**

Possibly.  Although the observations are independent of each other, the fact that the same professor appears in several of them might affect the linear regression.

### Exercise 18
**Based on your final model, describe the characteristics of a professor and course at University of Texas at Austin that would be associated with a high evaluation score.**

According to this formula, the highest score would be for:

- A Male professor

- Not a minority

- Received Education in English 

- Young

- A High percentage of students completed the evaluations

- Teaches a one credit course

- High beauty average

- Has a black and white picture

### Exercise 19
**Would you be comfortable generalizing your conclusions to apply to professors generally (at any university)? Why or why not?**

No.  Since this is not an experiment but an observational study and the evaluations were taken at a single university, we don't have a sample representative of all colleges around united states.  


...

